---
title: "Lab10_Dommeti.1"
author: "Suvan Dommeti"
date: "`r Sys.Date()`"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Example 1 - Hypothesis Testing

```{r}
# provided code
library(readr)
vanillin <- c(38,46,41,36,42,40,34,41,33,
  46,41,39,44,45,41,37,39,36,
  40,43,37,41,39,36,47,36,41,
  38,38,38,33,39,41,36,38,44,39)

mean(vanillin)

```

## Method 1: Preforming the Hypothesis Test using a Confidence Interval

Test $H_O:\mu = 42$ vs. $H_1:\mu \neq 42$ at significance level $\alpha$ = 0.01. In other words, we are going to construct a 99% confidence interval for $\mu$

```{r}
alpha <- 0.01
n <- length(vanillin)
xbar <- mean(vanillin)
sigma <- 3.52 #given

z <- qnorm(1-alpha/2)

ci.lower <- xbar - z*sigma/sqrt(n)
ci.upper <- xbar + z*sigma/sqrt(n)

cat("We are 99% confident that the true mean vannilan level of this batch is between", ci.lower, "and", ci.upper)

```


Since 42 is not contained within this confidence interval we reject the null hypothesis $H_O$, We conclude that the true mean vanilla level is statistically significantly different from 42 ppt.

## Method 2: Testing using Rejection Regions

Test $H_O:\mu = 42$ vs. $H_1:\mu \neq 42$ at significance level $\alpha$ = 0.01. We will being using a z test statistic. Under the null hypothesis, this test statistic $Z$ follows the distribution $Z\sim N(0,1)$

We reject the null hypothesis when the test statistic $Z_{obv}$ is either to high or too low

```{r}
alpha <- 0.01

rr.lower <- qnorm(alpha/2)
rr.upper <- qnorm(1-alpha/2)

cat("the rejection region is (", Inf, "to", rr.lower, "] and [", rr.upper, "to", Inf, ")")
```


We would reject the null hypothesis when our observed test statistic falls within either of these two regions. 

```{r}
z_obs <- (xbar - 42) / (sigma / sqrt(n))
```

We observe that $z_obs$ = -4.25. Since this falls within the rejection region, we must reject the null hypothesis $H_0$. We conclude the alternative hypotheses, which is that $\mu \neq 42$.
The true mean vannilan level is statistically significantly different than 42 ppt


## Method 3: Using the z-test function

```{r}
library(DescTools)

ZTest(vanillin, alternative = "two.sided", mu = 42, sd_pop = 3.52, conf.level = 1-alpha)

```

# Problem 2

```{r}

# provided parameters
mu0 <- 42 
alpha <- 0.01 

# calculate sample statistics
n <- length(vanillin)
xbar <- mean(vanillin)
s <- sd(vanillin)

# calculate t statistic
t_obs <- (xbar - mu0) / (s / sqrt(n))

# calculate critical value for two-sided test
t_critical <- qt(1 - alpha/2, df = n-1)

cat("Calculated t statistic (t_obs):", t_obs)
cat("\n", "Critical value for two-sided test (t_critical):", t_critical)

cat("\n")
# perform hypothesis test (this is a more automated way of doing it)
if (abs(t_obs) > t_critical) {
  cat("Reject the null hypothesis H0: mu = mu_0")
} else {
  cat("Fail to reject the null hypothesis H0: mu = mu_0")
}


```


# Problem 3

## part (a)

The test statistic we would use for this hypothesis test is the t-statistic, since the population variance is assumed to be unknown. The t-statistic is given by:

$t = \frac{xbar - \mu_0}{s / \sqrt{n}}$

where xbar is the sample mean, $\mu_0$ is the hypothesized population mean (in this case, $\mu_0$ = 8), s is the sample standard deviation, and n is the sample size (in this case, n = 30).

## part (b)

```{r}
# Given parameters
mu <- 8 # hypothesized population mean
sigma <- 1 # population standard deviation
n <- 30 # sample size
alpha <- 0.05 # significance level

# Calculate critical value for one-sided test
t_critical <- qt(alpha, df = n-1)

cat("Critical value for rejection region (t_critical):", t_critical)

```

## part (c)

```{r}
# Generate a large number of samples
n_samples <- 10000
test_statistics <- numeric(n_samples)

for (i in 1:n_samples) {
  # Generate sample data
  sample_data <- rnorm(n, mean = mu, sd = sigma)
  
  # Calculate sample statistics
  xbar <- mean(sample_data)
  s <- sd(sample_data)
  
  # Calculate test statistic
  test_statistics[i] <- (xbar - mu) / (s / sqrt(n))
}

```


## part(d)

```{r}
# Count proportion of test statistics in rejection region
proportion_rejection <- mean(test_statistics < t_critical)

cat("Proportion of test statistics in rejection region:", proportion_rejection)

cat("\n")


if (proportion_rejection > alpha) {
  cat(" The test statistics are falling in the rejection region more frequently than expected by chance, indicating rejection of the null hypothesis.")
} else {
  cat(" The test statistics are not falling in the rejection region more frequently than expected by chance, indicating failure to reject the null hypothesis.")
}

```


## part (e)

```{r}
# Plot histogram of test statistics
hist(test_statistics, freq = FALSE, main = "Sampling Distribution of Test Statistics",
     xlab = "Test Statistic", ylab = "Density", col = "mistyrose3")
lines(density(test_statistics), lwd = 2)
abline(v = t_critical, col = "red", lty = 2)
legend("topright", legend = c( "Critical Value"), lwd = c(2), col = c("red"), lty = c(2))

```

# Problem 4 (Using this one as notes/refernce for midterm)

```{r}
library(stats)


n <- 19                 # Sample size
mu <- 8                 # Population mean
sigma <- 1              # Population standard deviation
alpha <- 0.01           # Significance level

# Compute critical value
df <- n - 1             # Degrees of freedom for chi-squared distribution
chi_critical <- qchisq(1 - alpha, df)

# Generate samples and compute test statistic
num_samples <- 10000    # Number of samples for Monte Carlo simulation
test_statistics <- numeric(num_samples)  # Store test statistics

for (i in 1:num_samples) {
  # Generate sample from N(mu, sigma)
  sample <- rnorm(n, mean = mu, sd = sigma)
  
  # Compute sample variance
  sample_variance <- var(sample)
  
  # Compute test statistic
  test_statistic <- (n - 1) * sample_variance / sigma^2
  
  # Store test statistic
  test_statistics[i] <- test_statistic
}

# Compute proportion of test statistics in rejection region
proportion_rejection <- mean(test_statistics > chi_critical)

# Print results
cat("Rejection region proportion:", proportion_rejection, "\n")

# Plot sampling distribution
hist(test_statistics, freq = FALSE, col = "mistyrose3", border = "black",
     main = "Sampling Distribution of Test Statistics",
     xlab = "Test Statistic", ylab = "Density")
curve(dchisq(x, df), add = TRUE, col = "red", lwd = 2, n = 1000)
legend("topright", legend = c("Sampling Distribution", "Chi-squared Distribution"),
       col = c("mistyrose3", "red"), lwd = c(1, 2), bg = "white")

```

