---
title: "Lab 5 Dommeti.1"
author: "Suvan Dommeti"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

# Part (a)

$$L(\mu | x_1, x_2, ..., x_n, \sigma^2) = (2 ^ \pi)^{ -\frac{n}{2} } e^{ -\frac{1}{2}\sum_{i=1}^n (x_i - \mu)^2}$$

## Part (b)

```{r}
##make a function based of the math we did earlier
L.normal <- function(mu, obs.data) {
  n <- length(obs.data)
  
  likelihood <- (2*pi)^(-n/2) * exp(-1/2 * sum( (obs.data - mu)^2 ) )
  
  return(likelihood)
}

```

## Part (c)

```{r}
frogs <- c(4.57, 4.73, 2.24, 4.16, 4.53, 3.79, 2.48, 4.74, 3.18)

xbar <- mean(frogs)

print(xbar)
```

## Part (d)

```{r}
mu.domain <- seq(0, 7, by = 0.05)

likelihood.mu.storage <- c()

for(i in 1:length(mu.domain)){
  
  likelihood.mu.storage[i]  <- L.normal( mu.domain[i], obs.data = frogs)
  
}

plot(mu.domain, likelihood.mu.storage,
     type = "l",
     xlab = "mu",
     ylab = "L ( mu | x, sigma)",
     main = "Normal Likelihood of Frog Weights",
     col = "mistyrose3")

##Likelihood estimate is 3.8 
```

## Part (e)

```{r}
optimize(f = L.normal,
         interval = c(0,7),
         maximum = TRUE,
         obs.data = frogs)
```

## Part (f)

"\$maximum" is the value of $\mu$ that maximizes the likelihood function $L (mu | x, \sigma)$

"\$objective" is the value of the likelihood function $L (mu | x, \sigma)$ evaluated at $\mu = 3.824$

## Part (g)

```{r}
log.L.normal <-  function(mu, obs.data){
  
  n <- length(obs.data)
  
  likelihood <- (2*pi)^(-n/2) * exp(-1/2 * sum( (obs.data - mu)^2 ) )
  
  ##all we have to do is take the log at the end
  return(log(likelihood))
}

log.likelihood.mu.storage <- c()

for(i in 1:length(mu.domain)){
  
  log.likelihood.mu.storage[i]  <- log.L.normal( mu.domain[i], obs.data = frogs)
  
}

plot(mu.domain, log.likelihood.mu.storage,
     type = "l",
     xlab = "mu",
     ylab = "L ( mu | x, sigma)",
     main = "Log Likelihood of Frog Weights",
     col = "mistyrose3")
```

## Part (h)

```{r}
optimize(f = log.L.normal,
         interval = c(0,7),
         maximum = TRUE,
         obs.data = frogs)
```

## Part (i)

Because the log function is monotonically increasing

# Problem 2

## Part (a)

The likelihood function for a data set of exponentially distributed random variables with parameter $\lambda$ is:

$$L(\lambda|x_1, x_2, ..., x_n) = \lambda^n e^{-\lambda\sum_{i=1}^{n}x_i}$$

## Part (b)

```{r}
L.Exponential <- function(lambda, obs.data){
  n <- length(obs.data)
  
  likelihood <- lambda^n * exp(-lambda * sum(obs.data))
  
  return(likelihood)
}
```

## Part (c)

```{r}
 mercury <- c(0.51, 0.02, 0.15, 0.46, 0.11, 0.04, 0.39, 0.52, 0.2, 0.17, 0.01, 0.02, 0.32, 1.37)

lambda.MLE <- 1 / mean(mercury)

print(lambda.MLE)

```

## Part (d)

```{r}
lambda.domain <- seq(0, 10, by = 0.05)

likelihood.lambda.storage <- c()

for(i in 1:length(lambda.domain)){
  
  likelihood.lambda.storage[i] <- L.Exponential(lambda.domain[i], obs.data = mercury)
  
}

plot(lambda.domain, likelihood.lambda.storage,
     type = "l",
     xlab = "lambda",
     ylab = "L(lambda | x)",
     main = "Exponential Likelihood of Mercury Concentrations",
     col = "mistyrose3")
```

The likelihood function is uni-modal and peaks at around 3, which is consistent with the MLE computed in part (c).

## Part (e)

```{r}
optimize(f = L.Exponential,
         interval = c(0, 10),
         maximum = TRUE,
         obs.data = mercury)
```

## Part (f)

```{r}
log.L.Exponential <- function(lambda, obs.data){
  n <- length(obs.data)
  
  log.likelihood <- n*log(lambda) - lambda*sum(obs.data)
  
  return(log.likelihood)
}

log.likelihood.lambda.storage <- c()

for(i in 1:length(lambda.domain)){
  
  log.likelihood.lambda.storage[i] <- log.L.Exponential(lambda.domain[i], obs.data = mercury)
  
}

plot(lambda.domain, log.likelihood.lambda.storage,
     type = "l",
     xlab = "lambda",
     ylab = "log L(lambda | x)",
     main = "Log Likelihood of Mercury Concentrations",
     col = "mistyrose3")

```

The log-likelihood function is also unimodal and peaks at around 3, which is consistent with the MLE computed in part (c).

## Problem 3 

```{r}

river.data <- c("Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Flood", "Safe", "Safe", "Safe",
  "Flood", "Safe", "Safe", "Flood", "Flood", "Safe", "Safe", "Flood", "Safe", "Safe", "Safe", "Safe", "Safe",
  "Safe", "Safe", "Flood", "Safe")


likelihood <- function(p) {
  prod(ifelse(river.data == "Safe", p, 1-p))
}

p_seq <- seq(from = 0, to = 1, by = 0.01)
L <- sapply(p_seq, likelihood)
plot(p_seq, L, type = "l", xlab = "p", ylab = "Likelihood", main = "Likelihood function for p")




```

```{r}
MLE <- optimize(f = function(p) -likelihood(p), interval = c(0, 1), maximum = TRUE)
p_MLE <- MLE$maximum
p_MLE
```

The maximum likelihood estimate for p is approximately 0.767. This means that there is a 76.7% chance that the river is at a safe level on any given day. This estimate seems reasonable given that the river is at a safe level for the majority of the 30-day period.

```{r}
log_likelihood <- function(p) {
  sum(ifelse(river.data == "Safe", log(p), log(1-p)))
}
L_log <- sapply(p_seq, log_likelihood)
plot(p_seq, L_log, type = "l", xlab = "p", ylab = "Log-Likelihood", main = "Log-Likelihood function for p")


MLE_log <- optimize(f = function(p) -log_likelihood(p), interval = c(0, 1), maximum = TRUE)
p_MLE_log <- MLE_log$maximum
p_MLE_log

```

The maximum of the log-likelihood function occurs at the same value of p as the maximum of the likelihood function, which is approximately 0.767.
